The function-sheaf dictionary for smooth commutative group schemes over finite fields

### 2014-08-25 crelle-dictionary.tex
Working through paper, small changes

\begin{proposition}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
%We first show that $S_G$ is injective.  
Suppose $[\qcs{L}] \in \QCiso{G}$ with $S_G([\qcs{L}]) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\sigma$ for some $\sigma \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\sigma_x : \gqcs{L}_x \to \EEx$ by $v_x \mapsto \sigma(x)$.
Then the indexed set of isomorphisms $\{ \sigma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

To see that $S_G$ is surjective, begin with $\alpha\oplus\beta\oplus 0 \in Z^2(E^\bullet_G)$.
Since $d_{\Weil{}} \beta = 0$, we may define $a = \alpha \in C^2({\bar G},\EEx)$ and
$b = \beta(\Frob{}) \in C^1({\bar G},\EEx)$, which are related to $\alpha$ and $\beta$ as above.
Set $\gqcs{L}_x = \EEx$, define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes\gqcs{L}_y$
by $\mu_{x,y}(1) = a(x,y) 1\otimes 1$ and $\phi_x : \gqcs{L}_{\Frob{G}(x)} \to \gqcs{L}_x$ by $\phi_x(1)= b(x)$.
Then \ref{qc.1} holds since $d_G \alpha =0$ and \ref{qc.2} holds since $d_{\Weil{}}\alpha =d_G \beta$.
Tracing the construction backward, we have defined a quasicharacter sheaf $\qcs{L}$ on $G$ with
$S_G(\qcs{L}) = [\alpha\oplus\beta\oplus\gamma]$, showing that $S_G$ is surjective.
\end{proof}

@@

In this section we find the relationship between the group homomorphisms $\TrFrob{G}$ and $S_G$, and use this to find the group $\QCiso{G}$ when $G$ is an \'etale commutative group scheme over $\Fq$.

@@

\begin{lemma} \label{lem:dual-inv}
Let $X$ be an abelian group equipped with an action of $\Weil{}$.
 Then
\begin{align*}
 (X^*)_{\Frob{}} &\to (X^{\Frob{}})^* \\
 [f] &\mapsto f|_{X^{\Frob{}}}
\end{align*}
is an isomorphism.
\end{lemma}

\begin{proof}
We can describe $X^{\Frob{}}$ as the kernel of the map $X \xrightarrow{\Frob{}-1} X$;
let $Y = (\Frob{}-1)X$ be the augmentation ideal.  Dualizing the sequence
\[
 0 \to X^{\Frob{}} \to X \to Y \to 0
\]
yields
\[
 0 \to Y^* \to X^* \to (X^{\Frob{}})^* \to \Ext^1_\ZZ(Y, \EEx).
\]
Since $\Ext^1_\ZZ(-,\EEx)$ vanishes, we get a natural isomorphism from the cokernel of $Y^* \xrightarrow{\Frob{}-1} X^*$ to $(X^{\Frob{}})^*$.
\end{proof}

\begin{proposition}\label{prop:sur_etale}
If $G$ is \'etale then $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is surjective.
\end{proposition}
\begin{proof}
Pick $\chi \in G(\Fq)^*$.
By Lemma~\ref{lem:dual-inv}, there is some $b\in G(\bFq)^*$ such that $b\vert_{G(\Fq)^*} = \chi$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$.
Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = 1 \otimes 1$ and
$\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$ by $\phi_{x}(1) = b(x)$.
Since $b : G(\bFq) \to \EEx$ is a group homomorphism,
condition \eqref{nohom} is satisfied with $a =1$ and $\qcs{L} = (\gqcs{L}, \mu, \phi)$
is a quasicharacter sheaf with $\trFrob{\qcs{L}} = \chi$.
\end{proof}

>>>>

\begin{proposition}
If $G$ is \'etale then $S_G$ is an isomorphism.
\end{proposition}
\begin{proof}
%We first show that $S_G$ is injective.  
Suppose $[\qcs{L}] \in \QCiso{G}$ with $S_G([\qcs{L}]) = [\alpha \oplus \beta] = 0$,
so that $\alpha \oplus \beta = d\sigma$ for some $\sigma \in C^0(\Weil{},C^1({\bar G},\EEx) = C^1({\bar G},\EEx)$.
For each $x\in {\bar G}$, define $\sigma_x : \gqcs{L}_x \to \EE$ by $\sigma_x : v_x \mapsto \sigma(x)$.
Then the indexed set of isomorphisms $\{ \sigma_x : \gqcs{L}_x \to \EE \tq x\in {\bar G}\}$
defines an isomorphism $\qcs{L} \to (\EE)_G$.
Since $\qcs{L} = 0 \in \QCiso{G}$, $S_G$ is injective.

To see that $S_G$ is surjective, begin with $\alpha\oplus\beta\oplus 0 \in Z^2(E^\bullet_G)$.
Since $d_{\Weil{}} \beta = 0$, we may define $a = \alpha \in C^2({\bar G},\EEx)$ and
$b = \beta(\Frob{}) \in C^1({\bar G},\EEx)$, which are related to $\alpha$ and $\beta$ as above.
Set $\gqcs{L}_x = \EE$, define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes\gqcs{L}_y$
by $\mu_{x,y}(1) = a(x,y) (1\otimes 1)$ and $\phi_x : \gqcs{L}_{\Frob{G}(x)} \to \gqcs{L}_x$ by $\phi_x(1)= b(x)$.
Then \ref{qc.1} holds since $d_G \alpha =0$ and \ref{qc.2} holds since $d_{\Weil{}}\alpha =d_G \beta$.
Tracing the construction backward, we have defined a quasicharacter sheaf $\qcs{L}$ on $G$ with
$S_G(\qcs{L}) = [\alpha\oplus\beta\oplus 0]$, showing that $S_G$ is surjective.
\end{proof}

@@

In this section we fit the group homomorphisms $\TrFrob{G}$ and $S_G$ into a commutative diagram, determining the kernel and cokernel of the homomorphism $\TrFrob{G} : \QCiso{G} \to G(k)^*$ when $G$ is an \'etale commutative group scheme over $\Fq$.

@@

\begin{lemma} \label{lem:dual-inv}
Let $X$ be an abelian group equipped with an action of $\Weil{}$.
 Then
\begin{align*}
 (X^*)_{\Weil{}} &\to (X^{\Weil{}})^* \\
 [f] &\mapsto f|_{X^{\Weil{}}}
\end{align*}
is an isomorphism.
\end{lemma}

\begin{proof}
We can describe $X^{\Weil{}}$ as the kernel of the map $X \xrightarrow{\Frob{}-1} X$;
let $Y = (\Frob{}-1)X$ be the augmentation ideal.  Dualizing the sequence
\[
 0 \to X^{\Weil{}} \to X \to Y \to 0
\]
yields
\[
 0 \to Y^* \to X^* \to (X^{\Weil{}})^* \to \Ext^1_\ZZ(Y, \EEx).
\]
Since $\Ext^1_\ZZ(-,\EEx)$ vanishes, we get a natural isomorphism from the cokernel of $Y^* \xrightarrow{\Frob{}-1} X^*$ to $(X^{\Weil{}})^*$.
\end{proof}

\begin{proposition}\label{prop:sur_etale}
If $G$ is \'etale, then $\TrFrob{G} : \QCiso{G} \to G(\Fq)^*$ is surjective.
\end{proposition}
\begin{proof}
Pick $\chi \in G(\Fq)^*$.
By Lemma~\ref{lem:dual-inv}, there is some $b\in G(\bFq)^*$ such that $b\vert_{G(\Fq)} = \chi$.
Set $\gqcs{L}_x = \EE$ for every $x\in G(\bFq)$.
Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by $\mu_{x,y}(1) = 1 \otimes 1$ and
$\phi_{x} : \gqcs{L}_{\Frob{}(x)} \to \gqcs{L}_x$ by $\phi_{x}(1) = b(x)$.
Since $b : G(\bFq) \to \EEx$ is a group homomorphism,
condition \eqref{nohom} is satisfied with $a =1$ and $\qcs{L} = (\gqcs{L}, \mu, \phi)$
is a quasicharacter sheaf with $\trFrob{\qcs{L}} = \chi$.
\end{proof}
